Geometric variational problems

### 2020-11-13 NSF2020/Description/preliminarywork.tex
Update preliminarywork.tex

\begin{wraptable}[10]{r}{0.55\textwidth}
\begin{center}
  \begin{tabular}{|l|r|r|p{0.1\textwidth}|}
    \hline
          & HAP         & exper. value & ref.\\
    \hline 
    $\KB$ & $11 \pm 2.5$ \kBT & 10 \kBT                  & \cite{Naetal15,VeBrPa15,NAGLE2000159,PhysRevLett.113.248102}\\
    \hline 
    $\KA$  & $34$ \kBT nm$^{-2}$ & 30--40 \kBT nm$^{-2}$              & \cite{Nagle17, Nagle17-2}\\
    \hline 
    $\KTH$  & $12$ \kBT \; nm$^{-2}$ & 10 \kBT nm$^{-2}$  &  \cite{KUZMIN2005, KoNa15} \\ \hline
  \end{tabular}
\end{center}
\caption{\label{tab:moduli} \footnotesize Comparision of values elastic
  moduli from the experimental literature and values derived by HAP
  simulation.} 
\end{wraptable}
We performed three types of simulations. The first measured bending by loading
a planar bilayer. The second used a harmonic bond to dilate a circular bilayer.
The third measured tilt using a decay equation from  \cite{KUZMIN2005}.
These simulations isolated the deformations of \eqref{ansatz3}, enabling
us to read off elastic moduli from simulation data. 
The results agreed remarkably well with the values reported in the experimental literature, 
Table~\ref{tab:moduli}. The agreement is underscored by the fact
that the two, main HAP parameters $\gamma$ and $\rho$ were assigned
physical values from the outset, rather than being tuned.

\subsection{Two-dimensional vesicle hydrodynamics in shear flow} 
%The motion of vesicles in shear flow is an important problem in the
%applied mathematics because simulations can reveal mechanical
%properties of membranes and lead to an enhanced understanding of
%deformable particle laden flows \cite{Sinha15}. 
%
%
\begin{wrapfigure}[10]{r}{0.2\textwidth}
\centerline{\includegraphics[width=0.2\textwidth]{figures/PW_fig5.pdf}}
\vspace{-8pt}
\caption{\label{fig:rupture} \footnotesize Rupture of tank-treading vesicle under strong shear flow.}
\end{wrapfigure}
%
To implement a vesicle in shear flow in the context of hydrophobic
potentials and mobility problem, we consider a shear flow in the
far-field $\mathbf{u}_{\infty} = Uy\mathbf{i}_x$ in the direction of the
$x$-axis (Figure \ref{fig:tanktreading}A). As illustrated in Figure
\ref{fig:tanktreading}, the particle based approach supports
inter-leaflet slip, and this can be used to determine inter-leaflet and
in-plane shear viscosities. 
%
%This field satisfies the linear Stokes system but does not give rise to a rigid motion at the particle interfaces. 
%To have a rigid motion, we change variables $\mathbf{u} = \tilde{\mathbf{u}}+ \mathbf{u}_{\infty}$ and 
%for the new field $\tilde{\mathbf{u}}$ vanishing at infinity we let 
%$\tilde{\mathbf{u}}|_{\partial P_i} = \mathbf{v}_i + \boldsymbol{\omega}_i \times (\mathbf{x} - \mathbf{a}_i)$ 
%where $(\mathbf{v}_i,\boldsymbol{\omega}_i)$ are the unknown translation and angular velocities of the 
%$i$th particle $P_i.$  
%
%\begin{wrapfigure}[17]{l}{0.4\textwidth}
%\centerline{\includegraphics[width=0.4\textwidth]{figures/PW_fig2.pdf}}
%\caption{\label{fig:demixing} An initial assembly of small and 
%large particles spontaneous segregates into two smaller bodies. }
%\end{wrapfigure}
The HAP simulations show vesicle tank-treading. Under the external shear flow, the initially circular 
vesicle rotates in the clockwise direction. As the rate of rotation increases, the vesicle approaches
a steadily tank-treading ellipse. In Figure \ref{fig:tanktreading}B-D, the solid curves are ellipses fit to the particle centers
and midplane respectively. In the non-dimensionalized system, the particles have diameter 2, on the order of $\rho,$ 
and the vesicle diameter is about 14. 
%\todo[inline]{missing units. nm?}
Figure~\ref{fig:tanktreading}E shows the aspect ratio of the major to minor axes reaching an equilibrium value in the 
red and blue curves, yet oscillating in the high-shear rate (yellow) curve.
The tank-treading vesicle elongates and becomes more horizontal 
with an increase in flow rate or 
with a decrease in stiffness (effected by decreasing $\rho = 4$ to $\rho = 2$).

For large shear flow rates, there is an increase in arc length. Here arc
length refers to the the mid-plane circumference. Thus, some of the
external force is going into stretching the vesicle---the other part is
going into bending and viscous dissipation. From our experiments, we
find that the vesicle ruptures once stretching exceeds about 5 \% (see
Figure \ref{fig:rupture}). Finally, movies of the tank-treading motion
show a slip velocity between the outer and inner leaflets Figure
\ref{fig:tanktreading}G. We have illustrated this by tracking the
distance between two reference particles in the inner and outer leaflet
(Figure \ref{fig:tanktreading}B \& D, green and blue particles). With
moderate shear rates or greater adhesion, the particle pair moves in
tandem (in Figure \ref{fig:tanktreading}H, blue and red curves, their
distance is more or less constant). For a large shear rate, the
particle separates as the two leaflets slide against one another.

\begin{figure}
\begin{center}
\includegraphics[width=1\textwidth]{figures/PW_fig1A-D.pdf}
\includegraphics[width=1\textwidth]{figures/PW_fig1E-H.pdf}
\end{center}
\vspace{-0.3in}
\caption{\label{fig:tanktreading}\footnotesize (A) A vesicle formed by
  amphiphilic particles in shear flow, and the tank-treading motion
  (B)--(D). The separation of particle pairs in (B) and (C) illustrate
  inter-leaflet slip.  (E)--(G) Tank-treading reaches a steady state in
  elliptical aspect ratio, major-axis angle, and circumference.}
\end{figure}

>>>>

\begin{wraptable}[10]{r}{0.55\textwidth}
\begin{center}
  \begin{tabular}{|l|r|r|p{0.1\textwidth}|}
    \hline
          & HAP                     & exper. value          & ref.\\
    \hline 
    $\KB$ & $11 \pm 2.5$ \kBT       & 10 \kBT               & \cite{Naetal15,VeBrPa15,NAGLE2000159,PhysRevLett.113.248102}\\
    \hline 
    $\KA$ & $34$ \kBT \; nm$^{-2}$  & 30--40 \kBT nm$^{-2}$ & \cite{Nagle17, Nagle17-2}\\
    \hline 
    $\KTH$ & $12$ \kBT \; nm$^{-2}$ & 10 \kBT nm$^{-2}$     & \cite{KUZMIN2005, KoNa15} \\ \hline
  \end{tabular}
\end{center}
\caption{\label{tab:moduli} \footnotesize Comparision of values elastic
  moduli from the experimental literature and values derived by HAP
  simulation.} 
\end{wraptable}
We performed three types of simulations. The first measured bending by loading
a planar bilayer. The second used a harmonic bond to dilate a circular bilayer.
The third measured tilt using a decay equation from  \cite{KUZMIN2005}.
These simulations isolated the deformations of \eqref{ansatz3}, enabling
us to read off elastic moduli from simulation data. The results agreed remarkably 
well with the values reported in the experimental literature, 
Table~\ref{tab:moduli}. The agreement is underscored by the fact
that the two, main HAP parameters $\gamma$ and $\rho$ were assigned
physical values from the outset, rather than being tuned.

The PIs BQ, RR, YNY and Peter Fu initiated a collaboration in summer, 2020. 
The present proposal is the outgrowth of this collaboration. We have created
a compact numerical solver and included background flow in the simulations. 
The background flow enters by replacing the third equation in \eqref{eq:stokes} 
with the condition $(\mathbf{u} -\mathbf{u}_{\infty})(\mathbf{x}) 
\to \mathbf{0}$ as $|\mathbf{x}| \to \infty$. To incorporate the far-field flow, 
and using the representation 
\begin{align}
\label{PowerMiranda}
  {\bf u} = {\bf u}_{\infty} + K\boldsymbol{\eta} + 
    \sum_{i=1}^N S(\cdot, {\bf a}_i) {\bf F}_i + 
                 R(\cdot, {\bf a}_i) {\bf G}_i.
\end{align}
The symbols $S$ and $R$ are stokeslets and rotlets supported at the respective
particle centers \cite{leal_2007} and $K\boldsymbol{eta}$ is a layer potential for the
unknown density function $\boldsymbol{\eta}.$ The
representation~\eqref{PowerMiranda} automatically satisfies all
equations, with the exception of the rigid motion conditions. The rigid
motion conditions follow by requiring the viscous stress vanishes across
the particle boundaries.

\begin{wrapfigure}[13]{l}{0.30\textwidth}
\centerline{\includegraphics[width=0.30\textwidth]{figures/BIError.pdf}}
  \vspace{-8pt}
\caption{
\label{fig:bierror}  
\footnotesize The false color map shows how numerical quadrature of
  layer potentials loses accuracy near the particle boundaries.  The
  color bar is for $\log_{10}.$}
\end{wrapfigure}
\textbf{Novel reciprocal identities.} 
The force and torque formulas~\eqref{forceandtorque} require the
hydrophobic stress along the particle boundaries. Unfortunately,
standard quadrature formulas to estimate $\nabla u$ on the particle
boundaries introduces large amounts of quadrature error
(Figure~\ref{fig:bierror}). Therefore, it is useful to devise reciprocal
identities for the force and torque on body $i$ that does not
require integration along body $i$. One identity, that we have proved,
is
\begin{align}
    \label{eq:reciprocal}
{\bf F}_{\text{hydro},i} = \sum_{j \neq i} \int_{\partial P_i}[\boldsymbol{\sigma}_{ij} + \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S,\quad
{\bf G}_{\text{hydro},i} = \sum_{j \neq i} \int_{\partial P_i} ({\bf
  x}-\mathbf{a}_i) \times [\boldsymbol{\sigma}_{ij} +
  \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S, 
\end{align}
for $i=1,\ldots,N$. Here, $u_i$ is the solution of the screened Laplace
equation when only the contribution from particle $i$ is considered and
$\boldsymbol{\sigma}_{ij} = \rho^{-1} u_iu_j {\bf I} + \rho(\nabla u_i
\cdot \nabla u_j {\bf I} - 2 \nabla u_i \otimes \nabla u_j)$.
Equation~\eqref{eq:reciprocal} does not require knowledge of $\nabla u$
on the particle boundary, and this is enormously beneficial for
calculating force and torque.

\subsubsection{Two-dimensional vesicles in shear flow}
We will compare the behavior of our particle-based vesicles in Stokes
flow to well-established models that enforce area incompressibility and
volume conservation in vesicle membrane
dynamics~\cite{torres-sanchez_millan_arroyo_2019,
mahapatra_saintillan_rangamani_2020, Steigmann99, C6SM02452A}. Our
preliminary results show that the particle-based bilayers have the same
large area modulus as real lipid bilayer membranes, so we expect
realistic results from our flow simulations. Under moderate shear rates,
the vesicle elongates and takes on the tank-treading elliptical shape.
We can directly check for area compressibility conservation properties
from the particle simulation by tracking the change in area density of
the midplane as a function of shear flow rate, along with evaluating the
tangential divergence of the velocity.

Bilayer membranes have a small permeability to water
\cite{323e9a2f0c58487ea82518d7a1f96485}, and modelers often assume a
vesicle membrane that conserves volume. In our particle-based approach,
water motion across the bilayer is limited by the size of the gaps
between the particles, which is an artifact of coarse-graining.  Making
these gaps smaller comes at the expense of numerical accuracy, and we
will assess if there is a reasonable trade-off between approximate
volume conservation and simulation complexity.

Finally, the particle-based vesicle bilayers have two distinct leaflets
consisting of the particles in contact with the vesicle interior, and
those in contact with the surrounding fluid. Since there is a nonzero
separation between these layers, an important effect we observe is that
the leaflets slide against one another under shear flow.  This implies
that part of the viscous dissipation in the aqueous phase is enhanced by
intermonolayer friction~\cite{SHKULIPA2005823, ShkulipaThesis}.
Intermonolayer slip enters zero-thickness membrane models as a velocity
jump boundary condition and leads to multiple solution
states~\cite{schwalbe_vlahovska_miksis_2010}.  In the present proposal,
intermonolayer slip is an artifact of monolayer independence  and we can
control for slip as a function of shear rate, vesicle diameter, and
particle geometry.

%\subsection{Two-dimensional vesicle hydrodynamics in shear flow} 
%The motion of vesicles in shear flow is an important problem in the
%applied mathematics because simulations can reveal mechanical
%properties of membranes and lead to an enhanced understanding of
%deformable particle laden flows \cite{Sinha15}. 
%
%
%\begin{wrapfigure}[10]{r}{0.2\textwidth}
%\centerline{\includegraphics[width=0.2\textwidth]{figures/PW_fig5.pdf}}
%\vspace{-8pt}
%\caption{\label{fig:rupture} \footnotesize Rupture of tank-treading vesicle under strong shear flow.}
%\end{wrapfigure}
%
%To implement a vesicle in shear flow in the context of hydrophobic
%potentials and mobility problem, we consider a shear flow in the
%far-field $\mathbf{u}_{\infty} = Uy\mathbf{i}_x$ in the direction of the
%$x$-axis (Figure \ref{fig:tanktreading}A). As illustrated in Figure
%\ref{fig:tanktreading}, the particle based approach supports
%inter-leaflet slip, and this can be used to determine inter-leaflet and
%in-plane shear viscosities. 
%
%This field satisfies the linear Stokes system but does not give rise to a rigid motion at the particle interfaces. 
%To have a rigid motion, we change variables $\mathbf{u} = \tilde{\mathbf{u}}+ \mathbf{u}_{\infty}$ and 
%for the new field $\tilde{\mathbf{u}}$ vanishing at infinity we let 
%$\tilde{\mathbf{u}}|_{\partial P_i} = \mathbf{v}_i + \boldsymbol{\omega}_i \times (\mathbf{x} - \mathbf{a}_i)$ 
%where $(\mathbf{v}_i,\boldsymbol{\omega}_i)$ are the unknown translation and angular velocities of the 
%$i$th particle $P_i.$  
%
%\begin{wrapfigure}[17]{l}{0.4\textwidth}
%\centerline{\includegraphics[width=0.4\textwidth]{figures/PW_fig2.pdf}}
%\caption{\label{fig:demixing} An initial assembly of small and 
%large particles spontaneous segregates into two smaller bodies. }
%\end{wrapfigure}
%The HAP simulations show vesicle tank-treading. Under the external shear flow, the initially circular 
%vesicle rotates in the clockwise direction. As the rate of rotation increases, the vesicle approaches
%a steadily tank-treading ellipse. In Figure \ref{fig:tanktreading}B-D, the solid curves are ellipses fit to the particle centers
%and midplane respectively. In the non-dimensionalized system, the particles have diameter 2, on the order of $\rho,$ 
%and the vesicle diameter is about 14. 
%\todo[inline]{missing units. nm?}
%Figure~\ref{fig:tanktreading}E shows the aspect ratio of the major to minor axes reaching an equilibrium value in the 
%red and blue curves, yet oscillating in the high-shear rate (yellow) curve.
%The tank-treading vesicle elongates and becomes more horizontal 
%with an increase in flow rate or 
%with a decrease in stiffness (effected by decreasing $\rho = 4$ to $\rho = 2$).

%For large shear flow rates, there is an increase in arc length. Here arc
%length refers to the the mid-plane circumference. Thus, some of the
%external force is going into stretching the vesicle---the other part is
%going into bending and viscous dissipation. From our experiments, we
%find that the vesicle ruptures once stretching exceeds about 5 \% (see
%Figure \ref{fig:rupture}). Finally, movies of the tank-treading motion
%show a slip velocity between the outer and inner leaflets Figure
%\ref{fig:tanktreading}G. We have illustrated this by tracking the
%distance between two reference particles in the inner and outer leaflet
%(Figure \ref{fig:tanktreading}B \& D, green and blue particles). With
%moderate shear rates or greater adhesion, the particle pair moves in
%tandem (in Figure \ref{fig:tanktreading}H, blue and red curves, their
%distance is more or less constant). For a large shear rate, the
%particle separates as the two leaflets slide against one another.

%\begin{figure}
%\begin{center}
%\includegraphics[width=1\textwidth]{figures/PW_fig1A-D.pdf}
%\includegraphics[width=1\textwidth]{figures/PW_fig1E-H.pdf}
%\end{center}
%\vspace{-0.3in}
%\caption{\label{fig:tanktreading}\footnotesize (A) A vesicle formed by
%  amphiphilic particles in shear flow, and the tank-treading motion
%  (B)--(D). The separation of particle pairs in (B) and (C) illustrate
%  inter-leaflet slip.  (E)--(G) Tank-treading reaches a steady state in
%  elliptical aspect ratio, major-axis angle, and circumference.}
%\end{figure}
